---
title: "Tools for collaborating in teams"
subtitle: "(sample solution for the test case)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: "../extras/x-theme.css"
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height=5)
library(tidyverse)
```

## A test case

Using the gapminder data provided, we are asked to:

1. Visualize life expectancy over time for Canada in the 1950s and 1960s using a line plot.

2. Something is clearly wrong with this plot! Turns out there's a data error  in the data file: life expectancy for Canada in the year 1957 is coded as `999999`, it should actually be `69.96`. Make this correction.

3. Visualize life expectancy over time for Canada again, with the corrected data.

---
## Step-by-step plan of work

1. Read the data

2. Focus on values of Canada

3. Visualize the values for Canada

4. See whether we can find the problem

5. ... then we'll see ...

---
## Reading the data 

```{r}
life5060 <- read.csv("https://raw.githubusercontent.com/Stat480-at-ISU/materials/master/01_collaborative-environment/data/gapminder-5060.csv")
head(life5060)
```

---
## Focus on the values for Canada

```{r}
canada <- life5060 %>% filter(country == "Canada")
head(canada)
```

---
## Visualize 

Draw a line for the life expectancy in Canada
```{r}
canada %>% 
  ggplot(aes(x = year, y = lifeExp)) +
  geom_line()
```

---
## Find the problematic value and fix it

```{r}
canada %>% filter(year == 1957)

canada_fixed <- canada %>% mutate(
  lifeExp = replace(lifeExp, year==1957, 69.96)  
)
```

---
## visualize the fixed data

```{r}
canada_fixed %>% ggplot(aes(x = year, y = lifeExp)) +
  geom_line()
```

---
## ... on to the stretch goal ... 

Add lines for Mexico and the US

```{r}
canada_fixed %>% 
  ggplot(aes(x = year, y = lifeExp, group=country)) + geom_line() +
  geom_line(data = life5060 %>% 
              filter(country %in% c("United States", "Mexico")))
```

---
## ... on to the stretch goal ... 

Add lines for Mexico and the US, color lines by country and add a legend

```{r}
canada_fixed %>% 
  ggplot(aes(x = year, y = lifeExp, colour=country)) + geom_line() +
  geom_line(data = life5060 %>% 
              filter(country %in% c("United States", "Mexico")))

```

---
## How reproducible is this solution?

1. Navigate to https://github.com/stat480-at-isu/materials

2. Open the folder `01-collaborative_environment` 

3. Download the file `02-test-case_solution.Rmd`

3. Open the file in RStudio.

4. "Knit" the file.

---
## R Markdown

- code and text/documentation are interwoven: reproducible and self-documenting.

- extend or refine analyses by copying and modifying code blocks.

- disseminate your work by sharing the RMarkdown file
