---
title: "DATES & TIMES"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


### YOUR TURN #1

#### Q: Create date objects for today's date by typing the date in text format and converting it with one of the lubridate converter functions.

```{r}
library(lubridate)
ymd("2020-03-26")
mdy("03-26-2020")
```


#### Q: Try different formats of writing the date and compare the end results.

```{r}
ymd("2020-03-26", tz = "UTC")
```

#### Q: Use the appropriate lubridate function to parse each of the following dates:

```{r}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014
```

```{r}
mdy(d1)
ymd(d2)
dmy(d3)
mdy(d4)
mdy(d5)
```

### YOUR TURN #2

For this question use the `flights` data from the `nycflights13` package.

```{r}
library(nycflights13)
library(tidyverse)
```


#### Q: Use `make_datetime()` to create a date-time variable for `dep_time` and `arr_time`.   

**Hint**: use modular arithmetic `%/%` for hour and `%%` for minute.

```{r}
flights_dt <- flights %>% 
  filter(!is.na(dep_time), !is.na(arr_time)) %>% 
  mutate(dep_hour = dep_time %/% 100,
         dep_min = dep_time %% 100,
         dep_time = make_datetime(year, month, day, dep_hour, dep_min),
         arr_hour = arr_time %/% 100,
         arr_min = arr_time %% 100,
         arr_time = make_datetime(year, month, day, arr_hour, arr_min))
flights_dt
```

### YOUR TURN #3

For this question use the `flights_dt` data created in the last your turn.


#### Q: Use an accessor function to calculate the average departure delay by minute within the hour.  Use ggplot2 to plot your results.

```{r}
flights_dt %>% 
  mutate(minute = minute(dep_time)) %>% 
  group_by(minute) %>% 
  summarise(
    avg_delay = mean(arr_delay, na.rm = TRUE),
    n = n()) %>% 
  ggplot(aes(minute, avg_delay)) +
    geom_line()
```


### YOUR TURN #4

#### Q: Inspect the `budget` data set from the `classdata` package

```{r}
library(classdata)
budget %>% glimpse()
```


#### Q: Make sure the variable `Release Date` is a date format.

```{r}
summary(budget$ReleaseDate)
```

```{r}
budget %>% 
  arrange(desc(ReleaseDate)) %>%
  head()
```


#### Q: Plot a histogram of the variable

```{r}
budget %>% 
  ggplot(aes(x = ReleaseDate)) +
  geom_histogram(binwidth=365)
```


#### Q: Merge (`join`) budget and box office data (by movie name)  

```{r}
box_budget <- box %>% left_join(budget, by = "Movie")
head(box_budget)
```

#### Q: Is the time between the release of a movie and the date is equal to the number of weeks in theaters?

```{r}
box_budget %>% 
  filter(!is.na(ReleaseDate)) %>% 
  mutate(test_week = (ReleaseDate %--% Date) %/% dweeks(1)) %>% 
  select(Movie, ReleaseDate, Date, Week, test_week) %>% 
  head()

box_budget %>% 
  filter(!is.na(ReleaseDate)) %>% 
  mutate(test_week = ceiling((ReleaseDate %--% Date) / dweeks(1))) %>% 
  select(Movie, ReleaseDate, Date, Week, test_week) %>% 
  head()
```
