---
title: "TIME SERIES"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE, fig.height=3, fig.width=5
)
```


### YOUR TURN #1

#### Q: Load the `nasa` data from the package `GGally`

```{r}
library(tidyverse)
data(nasa, package="GGally")
```


#### Q:  for one location, draw a time line of Ozone over the time frame (`time`).  

```{r}
nasa %>% filter(id =="1-1") %>%
  ggplot(aes(x = time, y=temperature)) + 
  geom_point()
```


#### Q:  Plot separate lines for each of the years, i.e. put `month` on the x-axis and `ozone` on the y-axis for the same location. Is there a seasonal pattern apparent?  

```{r}
nasa %>% filter(id== "1-1") %>%
  ggplot(aes(x = month, y = ozone, group = year)) +
  geom_line(aes(colour = factor(year))) +
  geom_point()
```


#### Q:  Pick locations with x in 1:10 and y in 7:10. Plot temperature over time. Comment on the result.  

```{r}
nasa %>% 
  filter(x %in% 1:10, y %in% 7:10) %>%
  ggplot(aes(x = date, y=temperature, group=id)) +
  geom_line(alpha = .3) 
```


### YOUR TURN #2

#### Q: Load the box data from the package classdata

```{r}
data(box, package="classdata")
```

#### Q: For each movie and distributor, find:

(1) the highest total gross,

(2) the last date (and week) the movie was shown in theaters,

(3) the gross the movie made in the first week it was featured on the box office list

(4) the number of times the movie appears on the box office list

```{r}
box_summary <- box %>%
  group_by(Movie, Distributor) %>%
  summarize(
    maxGross = max(Total.Gross, na.rm=T),
    lastDate = max(Date, na.rm=T),
    lastWeek = max(Week, na.rm=T),
    minGross = min(Total.Gross, na.rm=T),
    n = n()
  )

box_summary %>% head()
```

