---
title: "MAPS"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


### YOUR TURN #1

```{r}
library(tidyverse)
```

#### Q: Use ggplot2 and pull out map data for all US counties: `counties <- map_data("county")`

```{r}
counties <- map_data("county")
```


#### Q: Draw a map of counties (polygons & path geom)

```{r}
counties %>% 
  ggplot(aes(x = long, y = lat)) + 
  geom_polygon(aes(group=group), size=.1, colour="white")
```


#### Q: Colour all counties called "story"

```{r}
counties %>% 
  ggplot(aes(x = long, y = lat)) + 
  geom_polygon(aes(group = group, fill = subregion == "story"), size = .1, colour = "white")
# 
# 
# counties %>% 
#   mutate(filled = subregion == "story") %>% 
#   ggplot(aes(x = long, y = lat)) + 
#   geom_polygon(aes(group = group, fill = filled), size = .1, colour = "white")
```


#### Q: Advanced: What county names are used most often?

Identify the counties: 

```{r}
counties %>% count(subregion) %>% arrange(desc(n)) # counts multiples

counties %>% 
  distinct(region, subregion) %>%
  count(subregion) %>% 
  arrange(desc(n))
```

Look at those countries on the map:

```{r}
counties %>% 
  ggplot(aes(x = long, y = lat)) + 
  geom_polygon(aes(group = group, fill= subregion == "jefferson"))

ggplot(counties, aes(x = long, y = lat)) + 
  geom_polygon(aes(group = group, fill = subregion == "washington"))

counties %>% 
  mutate(wash_jeff = replace(subregion, !(subregion %in% c("jefferson", "washington")), NA)) %>% # arrange(wash_jeff)
  ggplot(aes(x = long, y = lat)) + 
  geom_polygon(aes(group=group, fill= wash_jeff))
```


### YOUR TURN #2

For this question use `usa_sf()` from the `albersusa` package to retreive a U.S. state composite map as a shapefile. (run the code below)

```{r}
# devtools::install_github("hrbrmstr/albersusa")
library(albersusa)
states_sf <- usa_sf()
```

#### Q: Use geom_sf() and the states data to create a map of the US.


```{r}
library(sf)
ggplot(data = states_sf) + geom_sf()
```



### YOUR TURN #3

```{r}
data(fbi, package="classdata")
```

#### Q: Draw a choropleth map of the rate of motor vehicle thefts in 2012 across the US.

#### Q: `scale_fill_gradient2()` allows you to set a color scheme with two main colors. Read up on it and change the scheme in the first choropleth map.

```{r}
states <- map_data("state")
library(classdata)

fbi.map <- fbi %>% 
  filter(Year==2012) %>% 
  mutate(region = tolower(State), 
         rate = Count/Population*60000) %>% 
  left_join(states, by="region")

fbi.map %>% 
  filter(Type=="Motor.vehicle.theft") %>%
  ggplot(aes(x = long, y = lat)) +
  geom_polygon(aes(group = group, fill = rate))+
  scale_fill_gradient2(midpoint=200)
```



### YOUR TURN #4

```{r}
acc <- read.csv("https://raw.githubusercontent.com/DS202-at-ISU/labs/master/data/fars2016/accident.csv", stringsAsFactors = FALSE)
```


#### Q: Use the accident data to plot the geographic location of all accidents in the US in 2016.

#### Q: Plot accidents on a map of the US (use the map of the US as first layer)

```{r}
states %>% 
  ggplot(aes(x = long, y=lat)) +
  geom_path(aes(group=group)) +
  geom_point(data=filter(acc, between(LONGITUD, -130, 700)), 
             aes(x = LONGITUD, y=LATITUDE), 
             size = .1, alpha=.1, color = "violetred4"
             ) 
```


#### Q: Why would it be tricky to plot a choropleth map of the number of accidents by state?



### YOUR TURN #5


#### Q: The numbers for each state (`STATE`) are so-called fips codes.


#### Q: Sketch out the steps necessary to draw a choropleth map of the rate of fatal accidents by state.


```{r eval = FALSE}
states_sf <- usa_sf()
states_modified <- states %>% 
  mutate(STATE = as.numeric(as.character(states$fips_state)))
acc_mod <- acc %>% 
  group_by(STATE) %>% 
  mutate(count = n()) %>% 
  left_join(states_modified, by="STATE")

acc_mod %>% 
  #filter(between(LONGITUD, -130, 700)) %>% 
  ggplot() +
  geom_sf(aes(fill = count)) 

```

