---
title: "Coding style `r emo::ji('dancer')`"
ratio: 16x10
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: "../extras/x-theme.css"
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, echo = TRUE)
#xaringan::inf_mr(cast_from = '..') 
# install_github("hadley/emo")
library(emo)
```

## Style guide

<br/>
>"Good coding style is like correct punctuation: you can manage without it, butitsuremakesthingseasiertoread."
>
>Hadley Wickham

<br/><br/>
Style guide for this course is based on the [Tidyverse style guide](http://style.tidyverse.org/)

---
## File names & code chunk labels

Guidelines:

- Do not use spaces in file names, use `-` or `_` to separate words

- Use all lowercase letters

Example:

```{r eval = FALSE}
# Good
ucb-admit.csv

# Not so good
UCB Admit.csv
```

---
## Object names

Guidelines:

- Use `_` to separate words in object names
- Use informative but short object names
- Do not reuse object names within an analysis

Example:

```{r eval = FALSE}
# Good
acs_employed

# Not so good
acs.employed
acs2
acs_subset
acs_subsetted_for_males
```

---
## Spacing

Guidelines:

- Put a space before and after all infix operators (=, +, -, <-, etc.), and when naming arguments in function calls. 
- Always put a space after a comma, and never before (just like in regular English).

Example:

```{r eval = FALSE}
# Good
average <- mean(feet / 12 + inches, na.rm = TRUE)

# Not so good
average<-mean(feet/12+inches,na.rm=TRUE)
```

---

## ggplot

Guidelines:

- Always end a line with `+`
- Always indent the next line

Example:

```{r eval = FALSE}
# Good
ggplot(diamonds, mapping = aes(x = price)) +
  geom_histogram()

# Not so good
ggplot(diamonds,mapping=aes(x=price))+geom_histogram()
```

---

## Long lines

Guidelines:

- Limit your code to 80 characters per line. This fits comfortably <br>on a printed page with a reasonably sized font.
- Take advantage of RStudio editor's auto formatting for indentation <br>at line breaks.

---

## Assignment

Guidelines:

- Use `<-` not `=`

Example:

```{r eval = FALSE}
# Good
x <- 2

# Not so good
x = 2
```

---

## Quotes

Guidelines:

- Use `"`, not `'`, for quoting text. 

- The only exception is when the text already contains double quotes and no single quotes.

Example:

```{r eval = FALSE}
ggplot(diamonds, mapping = aes(x = price)) +
  geom_histogram() +
  # Good
  labs(title = "`Shine bright like a diamond`",
  # Good
       x = "Diamond prices",
  # Not so good
       y = 'Frequency')
```


---
## Resources

- Tidyverse style guide: http://style.tidyverse.org/

- Slides from datasciencebox.org
