---
title: "dplyr examples: <br>`group_by()` & `mutate()`"
ratio: 16x10
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: "../extras/x-theme.css"
    nature:
      highlightStyle: tomorrow
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, echo = TRUE)
#xaringan::inf_mr(cast_from = '..') 
library(tidyverse)
```

## Working with `fbi`

`fbi` data included in the `classdata` package

install the package: `devtools::install_github("haleyjeppson/classdata")`

```{r, message=FALSE}
library(tidyverse)
library(classdata)
data("fbi", package="classdata")
head(fbi)
```


---
class: yourturn
# Your turn 

For this your turn use the `fbi` data from the `classdata` package 

- Use `mutate()` to introduce a variable Rate into the fbi data

- Use `mutate()` to reorder the levels of types of crimes by (median) rate.

- Plot crime rates by type in side-by-side boxplots. Medians of the boxplots should be ordered.



```{r eval=FALSE, include=FALSE}
fbi <- fbi %>% group_by(Type)
fbi <- fbi %>% ungroup() %>%   mutate(
  Type = reorder(Type, Rate, FUN=median, na.rm=TRUE)
)
fbi %>% 
  ggplot(aes(x = Type, y = Rate)) +
  geom_boxplot() + coord_flip()
```

---
##  `group_by()` & `mutate()` 

```{r}
fbi %>% 
  group_by(Type) %>% 
  mutate(
    Rate = Count/Population*70000,
    Rank = rank(Rate) # ranks from lowest rate to highest rate
  ) %>% 
  dplyr::filter(Rank == 1) %>% 
  select(Type, State, Year, Rate)
```

---
class: yourturn
#  Your turn: <br>`group_by()` & `mutate()`

<br>
For this your turn use the `fbi` data from the `classdata` package 

- Introduce a ranking by rate for states for each type of crime and year.

- Focus on the top states. Find a visual that shows how often each state managed to take the top spot since 1961.

```{r eval=FALSE, include=FALSE}
fbi %>% 
  group_by(Type, Year) %>%
  mutate(
    Rate = Count/Population*70000,
    Rank = rank(Rate) # ranks from lowest rate to highest rate
  ) %>% 
  filter(Rank == 1) %>% 
  select(Type, State, Abb, Year, Rate) %>%
  ggplot(aes(x = Abb)) + 
  facet_wrap(~Type, scales="free", ncol=4) +
  geom_bar()
```
