---
title: "dplyr examples: <br>`group_by()` & `summarise()`"
ratio: 16x10
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: "../extras/x-theme.css"
    nature:
      highlightStyle: tomorrow
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, echo = TRUE)
#xaringan::inf_mr(cast_from = '..') 
library(tidyverse)
```

## The Happy data from GSS

The General Social Survey (GSS) has been run by NORC every other year since 1972 to keep track of current opinions across the United States. 

An excerpt of the GSS data called `happy` is available from the `classdata` package:

```{r, eval=FALSE}
devtools::install_github("haleyjeppson/classdata")
```

You can find a codebook with explanations for each of the variables at https://gssdataexplorer.norc.org/

---
## A first look

How many variables and how many observations does the data have? What do the variables mean?

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(classdata)
data("happy", package="classdata")
happy %>% str()
```

---
## The variable `happy`

```{r fig.height=5}
happy %>% 
  ggplot(aes(x = happy)) + geom_bar()
```

---
## The variable `age`

```{r, warning=FALSE, fig.height=5}
happy %>% ggplot(aes(x = age)) + geom_histogram(binwidth=1)
```

---
## The variable `degree`

```{r fig.height=5}
happy %>% ggplot(aes(x = degree)) + geom_bar()
```

---
## Summarising happiness

Use scores for `happy` factor to summarise overall happiness level,
- i.e. `not too happy` = 1, `pretty happy` = 2, and `very happy` = 3

```{r}
happy %>% summarise(
  m.happy = mean(as.numeric(happy), na.rm=TRUE)
  )
```
```{r}
happy %>% group_by(sex) %>% summarise(
  m.happy = mean(as.numeric(happy), na.rm=TRUE)
  )
```

---
class: yourturn
.center[
## Your turn: <br> `group_by()` & `summarise()`
]

For this your turn use the `happy` data from the `classdata` package 

- How do average happiness levels change over age?

- Are Republicans or Democrats happier? Compare average happiness levels over `partyid`.

- How are financial relations associated with average happiness levels? Is this association different for men and women?

- Find a plot that shows the differences for each one of the summaries.

---
class: yourturn
.center[
## Your turn: <br>asking questions 
]

For this your turn use the `happy` data from the `classdata` package 
- Investigate what other variable(s) might be associated with happiness. Plot it.

---
## Helper functions (1)

- `n()` provides the number of rows of a subset:

```{r}
happy %>% group_by(sex) %>% summarise(n = n())
```

- `tally()` is a combination of `summarise()` and `n()`

```{r}
happy %>% group_by(sex) %>% tally()
```

---
## Helper functions (2)

Count observations in groups

- `count()` is a further shortcut of `group_by()` and `tally()`:

```{r}
happy %>% count(sex, degree)
```

---
## Grouping & ungrouping

- `ungroup()` removes a grouping structure from a data set  

- necessary to make changes to a grouping variable (such as re-ordering or re-labelling)
