---
title: "Working with <br> missing values"
ratio: 16x10
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: "../extras/x-theme.css"
    nature:
      highlightStyle: tomorrow
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, echo = TRUE)
#xaringan::inf_mr(cast_from = '..') 
library(tidyverse)
```

## Missing values in R

- R allows consistent handling of mising values

- `NA` is the special code for "not available"

- `NaN` is the code for "not a number",  
  - e.g. in 0/0

- missing values propagate in calculus,  
  - e.g. for any object `x` we get `NA + x = NA`, `NA * x = NA`

---
## Essential functions 

Direct testing for missing values is resulting in an `NA`:

```{r}
x <- c(1, NA)
x == NA
```

</br>
Instead use function `is.na()` for vector `x`:

```{r}
is.na(x)
```

</br>
`complete.cases()` does the same for a data.frame

---
## Missing values essentials (2)

- **DANGER ZONE**: `na.omit()` removes all instances of mising values in an object (all rows with any missing value in case `x` is a data frame)

- Many functions have parameter `na.rm`

---
class: yourturn
# Your turn

Use the `box` data from the package `classdata`

- Are there any missing values in the dataset `box`?

- What are the values of `Rank` when `Rank.Last.Week` is missing?

- What is the dimension of the data set `box`, when removing all missing values with the function `na.omit()`?

- Why does the following statement fail? <br>
`box$Rank.Last.Week <- na.omit(box$Rank.Last.Week)` 


